---
title: "DAS-group-17"
format: pdf
editor: visual
---

## Import Data

```{r}
data = read.csv("dataset17.csv")
any(is.na(data))
```

## Library the Packages

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(gridExtra)
library(MASS)
library(pscl)
library(survival)
library(fitdistrplus)
library(gt)
```

## Introduction

The data we obtained was a part of data from the Dallas animal shelter data. The data given included 1678 observations including the variables with 6 category variables and 1 numerical variable.

Animal_type: 5 types of animals in the shelter Month: the month when the animal was admitted to the shelter. Year: the year when the animal was admitted to the shelter Intake_type: 3 different reasons for why the animal comes to the shelter. Outcome_type: 5 different results after the animal has finished assistance in the shelter. Chip_status: The status for the animal's chip status Time_at_shelter: The time for the animal to stay in the shelter?

In particular, this report presents numerical and graphical summaries of the data from the shelter for 7 different variables. A general linear model would be also fitted for the data, judging by some criteria statistics.

## Explanatory Data analysis

At the beginning of the EDA, we found that the year and month should be a category variable, so we need to change them first.

```{r}
data$month = factor(data$month)
data$year = factor(data$year)
```

### Summary Information

As the only numerical variables in this data, time_at_shelter should be consider with some summary statistics.

```{r}
#| label: summary statistics for time_at_shelter
summary1 = data%>%  
  summarize('Mean' = mean(time_at_shelter),
  'Median' = median(time_at_shelter),
  'St.Dev' = sd(time_at_shelter))%>%
mutate(Variable = "time_at_shelter")
summary2 = data%>%
  summarize(
            'IQR' = quantile(time_at_shelter,0.75)-quantile(time_at_shelter,0.25),
            'Min' = min(time_at_shelter),
            'Max' = max(time_at_shelter))%>%
mutate(Variable = "time_at_shelter")
```

```{r}
#| label: tbl-summarize_the_information_from_time_at_shelter_1
#| tbl-cap: Summary information of time_at_shelter part 1
#| warning: false
summary1%>%
  dplyr::select(Variable,everything())%>%
  gt()%>%
  fmt_number(columns = c(Mean, Median, St.Dev), decimals = 2) %>%
tab_style(style = cell_text(weight = "bold"),
locations = cells_column_labels())
```

```{r}
#| label: tbl-summarize_the_information_from_time_at_shelter_2
#| tbl-cap: Summary information of time_at_shelter part 2
#| warning: false
summary2%>%
  dplyr::select(Variable,everything())%>%
  gt()%>%
  fmt_number( decimals = 2) %>%
tab_style(style = cell_text(weight = "bold"),
locations = cells_column_labels())
  
```

According to @tbl-summarize_the_information_from_time_at_shelter_2, we can find that the mean of the time_at_shelter is not equal to median, the variance of time_at_shelter is much larger than the mean. This situation should be consider more in the part of formal analysis.

### Graphical Summaries

```{r}
#| label: fig-Histogram_for_time_at_shelter
#| fig-cap: Frequency Distribution for time at shelter
ggplot(data,aes(x = time_at_shelter))+
  geom_histogram(col = "white",fill = "steelblue")
```

From the @fig-Histogram_for_time_at_shelter and @tbl-summarize_the_information_from_time_at_shelter_2, the distribution for time_at_shelter should not be the poisson distribution. With data being overly dispersed, we could consider the time_at_shelter as a negative binomial distribution.

Based on data, it can be observed that the remaining six variables are categorical variables. Therefore, we can examine the distribution of categories by plotting bar plots.

```{r}
#| label: fig-bar_plot_for_categorical_variables
#| fig-cap: bar plots for categorical variables
a = ggplot(data,aes(x = animal_type))+
  geom_bar(fill = "steelblue")
b = ggplot(data,aes(x = month))+
  geom_bar(fill = "steelblue")
c = ggplot(data,aes(x = year))+
  geom_bar(fill = "steelblue")
d = ggplot(data,aes(x = intake_type))+
  geom_bar(fill = "steelblue")
e = ggplot(data,aes(x = outcome_type))+
  geom_bar(fill = "steelblue")
f = ggplot(data,aes(x = chip_status))+
  geom_bar(fill = "steelblue")
grid.arrange(a,b,c,d,e,f,nrow = 3)
```

The first figure in @fig-bar_plot_for_categorical_variables is a bar plot for the variable called animal_type, which shows that almost all the observations are cat and dog. With a similar situation, In the variable called outcome_type, the majority of the data are concentrated in three categories: adoption, euthanasia, and return to owner, with very few observations in the other two categories. We could also find the same situation in the bar plot for chip_status and intake_type, where only a very small portion of the data falls into the "CONFISCATED" and "UNABLE TO SCAN" category.

In the bar plot for months, we find that the data for June, July, and August are more abundant compared to other months.

Besides, the number of observations from 2016 is much less than the number of observations from 2017.

```{r}
#| label: fig-box_plot_for_categorical_variables
#| fig-cap: box plots for categorical variables
boxplot1 = ggplot(data,aes(x = animal_type,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
boxplot2 = ggplot(data,aes(x = month,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
boxplot3 = ggplot(data,aes(x = year,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
boxplot4 = ggplot(data,aes(x = intake_type,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
boxplot5 = ggplot(data,aes(x = outcome_type,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
boxplot6 = ggplot(data,aes(x = chip_status,y = time_at_shelter))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,20))
grid.arrange(boxplot1,boxplot2,boxplot3,boxplot4,boxplot5,boxplot6,nrow = 3)
```

From @fig-box_plot_for_categorical_variables, with a similar situation, we can see the means of the various categories are not the same in the variables of animal_type,intake_type, outcome_type, and chip_status.Conversely, the means of the time-based categorical variables are almost identical.
